---
title: "Data1 Analysis"
author: "wudayu PB19030934"
date: "12/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read data1
```{r}
data1<-read.csv("~/Desktop/data1.csv")
head(data1,10)
```
## Trail
It starts at (122.0951,	40.20291)	and goes northwestern reaching (122.0711,	40.21057) before gets back.

The destinatoin locates at (122.1018,	40.17257).
```{r}
longitude<-data1$LONG
latitude<-data1$LAT
plot(x=longitude,y=latitude,title("Trail of the car"))
```
## Descriptive Statistics
```{r}
summary(data1)
```


## Distributions of numerics
```{r}
consumption<-data1$CONSUM
acceleration<-data1$LONGACC
speed<-data1$SPEED
wheel<-data1$WHEEL
distance<-data1$CUMDIS
hist(consumption, main = paste("Histogram of consumption"))
hist(acceleration, main = paste("Histogram of acceleration"))
hist(speed, main = paste("Histogram of speed"))
hist(wheel, main = paste("Histogram of wheel"))
hist(distance, main = paste("Histogram of distance"))
```
## Explanation of histogram

1. Firstly, note that if we ignore the relatively small data (first bar in each histogram) in CONSUM, SPEED and WHEEL, it seems we'll have a more norm like distribution. This implies that sometimes the driving speed is low while in other times the speed has a norm-like distribution. It could be easily explained if we consider this scenario, at crosses and turnings, the speed slows down to a level. And if we don't consider these slowing down, the speed behaves like many other variables largely holding a norm distribution.

2. The Histogram of distance seems quite abnormal because that if we have a continuous stable record of one certain car, then it would seem like a uniform distribution.

However, since data is collected on a 15 seconds basis, the more one certain number occurred in cumulative distance, the longer the driver stayed in that road section, which leads to a low average speed in that place.

Then we may conclude that the driver followed such a driving history:

At the beginning, the car might be around certain district that only allows low speed driving, such as business circle, busy main road, etc.

Then the car sped up till entered another district with limitation.

After that the car might get on the highway and keep going with a relatively high speed.

Suddenly, it got stuck at around distance 3138 for a long time, which might be an accidence. 

Finally, it continue to go.

## Correlation
```{r}
M<-cor(data1[-1])
M
library(Hmisc)
rcorr(as.matrix(data1[-1]))
library(corrplot)
corrplot(method="square",M)
```

1. Clearly, distance, longitude and latitude are all closely related. Since the car goes largely towards southeast, distance certainly would have separately a positive and negative relationship with longitude and latitude.

2. Note that consumption is closely related with acceleration and speed, while acceleration is related with wheel. 

3. Taking the road condition into consideration, speed and consumption have something to do with longitude and latitude.

4. It's interesting to find that speed and wheel are closely related.

## Stationarity Analysis

When it comes to the driving behavior, stationary is most important thing we would consider first.

To picture the stationary of this drive, we ought to analysis variable speed using timeseries techniques.
```{r}
speedts<-ts(speed, start=1, end=1575)
plot(speedts,type='l',main='Speed against Time')

library(forecast)
library(tseries)
acf(speedts)
pacf(speedts)
Box.test(speedts,type = "Ljung-Box") 
ndiffs(speedts)
dspeedts <- diff(speedts)
plot(dspeedts)
acf(dspeedts)
pacf(dspeedts)
Box.test(dspeedts,type = "Ljung-Box") 
```

From the ACF and PACF graph, we find that ARIMA(1,1,3) is what we need for speedts.

To double check this result, we then apply auto.arima to this data and the results support our conclusion above.

```{r}
fit.speed<-auto.arima(speedts)
summary(fit.speed)
checkresiduals(fit.speed)
qqnorm(fit.speed$residuals)
qqline(fit.speed$residuals)
```
As the model settled down, some simple predictions are made as below.

```{r}
forecast(fit.speed,10,level=c(99.5))
plot(forecast(fit.speed,60))
```
Well, it seems that the party is over and no one would be on the raod :)

## Stationary Analysis 2
Apart from stationary of speed we get as above , LONGACC is a similar index for stationary of machine. 

We wonder whether these two characters will behave similarly. Or namely, ARIMZ(1,0,3) for LONGACC?

```{r}
accts<-ts(acceleration,start=1,end=1575,frequency = 60)
plot(accts)
acf(accts)
pacf(accts)
Box.test(accts,type = "Ljung-Box") 
ndiffs(accts)
fit.accts<-arima(accts,order=c(1,0,3))
checkresiduals(fit.accts)
```
The ACF graph shows that ARIMA(1,0,3) is not suitable for acceleration. 

This implies that acceleration get form sensor do not coordinate with acceleration get from speed. 

BUT WHY?

We think that LONGACC may be a acceleration along longitude thus shows a regular pattern and do not fit ARIMA model constructed on speed.

## Consumption Model
Consumption is very important for both driver and supplier.

From the correlation matrix and significance, we note that all factors are significantly (linearly) related with CONSUM.

Thus we now consider to predict CONSUM with other factors.

Since in this data set, outliers would greatly influence the performance of OLS and we apply ridge regression here.

```{r}
library(glmnet)
y<-scale(consumption)
x<-as.matrix(cbind(scale(longitude),scale(latitude),scale(speed),scale(wheel),scale(acceleration),scale(distance)))
model <- glmnet(x, y, alpha = 0)
cv_model <- cv.glmnet(x, y, alpha = 0)
best_lambda <- cv_model$lambda.min
best_lambda
plot(cv_model) 
best_model <- glmnet(x, y, alpha = 0, lambda = best_lambda)
coef(best_model)
plot(model, xvar = "lambda")
```
With this model we may help to predict consumption of cars around this area, which would certainly be beneficial for local gas stations.

## Other things we may include in driving behavior

One thing should be mention: the number of sudden acceleration. 

If a driver undergoes to many sudden accelerations, then the road condition can't be very good or the driver can't be very skillful.

```{r}
plot(dspeedts>10)
plot(dspeedts>7)
plot(dspeedts>5)
plot(dspeedts>3)
```




## To Be Continued
To better value and test our models, we then apply and develop them in data2